Understanding Implications of Dataset Choice for Feature Effect Estimation: A Simulation-Based Investigation through Error Decomposition

### 2025-02-21 paper/paper.tex

%N.B.: Author information (both in the \author{} and \authorrunning{} command) should only be present in the Camera-Ready Version of your paper. The version that you initially submit for review, ought to be double-blind. So, when initially submitting your paper, use:
%\author{Author information scrubbed for double-blind reviewing}
\author{Timo Heiß\inst{1}}
% You may leave out the orcidID information, if you want to.
% Use \corr to indicate the corresponding author. Note the spacing around the \corr command. Only one author can be the corresponding author.

@@

    \keywords{Interpretable Machine Learning  \and Feature effects \and Partial dependence plot \and Accumulated local effects}
\end{abstract}

@@

The optimal hyperparameters (OT) were chosen by tuning the models on a separate
data sample. Training and validation data were sampled independently from the
correspondings DGPs, with a training size of $n_{train}$ and a validation size
of 10000 to get a reliable performance estimate. While this scenario is
unrealistic in practice, it allows us to consider hyperparameters as
``pre-selected'' and avoid costly nested resampling strategies for the
simulation study. Each model was tuned for 200 trials using a
\textit{Tree-structured Parzen Estimator (TPE)}~\cite{bergstra2011algorithms}
using validation MSE as objective to minimize (equivalent to maximizing
R2-score).

>>>>

%N.B.: Author information (both in the \author{} and \authorrunning{} command) should only be present in the Camera-Ready Version of your paper. The version that you initially submit for review, ought to be double-blind. So, when initially submitting your paper, use:
%\author{Author information scrubbed for double-blind reviewing}
\author{Timo Heiß} % \inst{1}
% You may leave out the orcidID information, if you want to.
% Use \corr to indicate the corresponding author. Note the spacing around the \corr command. Only one author can be the corresponding author.

@@

    \keywords{Interpretable Machine Learning  \and Feature Effects \and Partial Dependence Plot \and Accumulated Local Effects}
\end{abstract}

@@

The optimal hyperparameters (OT) were chosen by tuning the models on a separate
data sample. Training and validation data were sampled independently from the
correspondings DGPs, with a training size of $n_{train}$ and a validation size
of 10000 to get a reliable performance estimate. While this scenario is
unrealistic in practice, it allows us to consider hyperparameters as
``pre-selected'' and avoid costly nested resampling strategies for the
simulation study. Each model was tuned for 200 trials using a
\textit{Tree-structured Parzen Estimator (TPE)}~\cite{bergstra2011algorithms}
using validation MSE as objective to minimize (equivalent to maximizing
R2-score). Hyperparameter search spaces were chosen following the
recommendations in~\cite{probst_tunability_2019}.
